New Bayes covariant coupling methods to cope with change of priors in one-vs-one classification

### 2025-06-26 coupling.tex
Fix cite

A notable example which uses this principle is the coupling method introduced in [\cite{hastie1998classification}]. Their method uses Kullback-Leibler divergence $\divm_\textrm{KL}$
\begin{align*}
	\divm\nolimits_\textrm{KL} (\boldsymbol{p}, \boldsymbol{q})= \sum p_i \log (p_i/ q_i)
\end{align*}

Kullback-Leibler divergence strongly penalizes  mispredictions when the true class is predicted to have very low probability. In the context of classification, when we are primarily concerned with accuracy , one could advantageously use other Bregman divergence derived from a proper score [\cite{gneiting2007strictly, buja2005loss}].

For instance, a popular alternative is divergence arising from the quadratic  score [\cite{gneiting2007strictly}] (also known as Brier score [\cite{brier1950verification}]). This divergence takes the form
$$
\divm\nolimits_\textrm{quad} (\boldsymbol{p}, \boldsymbol{q})= \sum (p_i - q_i)^2.
$$
Unlike Kullback-Leibler divergence, the penalty for incorrect prediction with arbitrarily low probability is bounded (by one). This quadratic score is the basis of the popular Wu-Lin-Weng's method [\cite{wu2004probability}].

@@

We note that the method is symmetric. The choice of quadratic scoring rule can also be viewed as canonical in view of axiomatic characterization given in  [\cite{selten1998axiomatic}].  However the choice of weights $w_{ij}$ to make the method computationally amenable is somewhat arbitrary. Thus we do not view Wu-Lin-Weng's method as canonical.

@@

As an example, consider the classification problem with three classes. We have three arboreal classifiers $\boldsymbol{s}_1, \boldsymbol{s}_2, \boldsymbol{s}_3$ which correspond to star graphs centered respectively at classes $C_1, C_2, C_3$. The classifier 
\begin{align}
\boldsymbol{e}_3 \stackrel{def}{=} \frac13 \boldsymbol{s}_1 \oplus \frac13 \boldsymbol{s}_2 \oplus \frac 13 \boldsymbol{s}_3  \label{eq:bc1}
\end{align}
is Bayes covariant \emph{and} symmetric. A key result proved in [\cite{vsuch2016bayes}] is that it is a unique  classifier having these two properties.

@@

\begin{table}[!ht]
\begin{tabular}{cm{2.5cm}m{1.5cm}m{1.5cm}ccm{1.5cm}m{1.5cm}}
&property & Hastie-Tibshirani & Wu-Lin-Weng & radial & normal & arboreal & Hinton's oracle \\
\hline 
\multirow{5}{*}{\begin{turn}{90}\makecell{exact}\end{turn}}
&uniqueness &  yes & yes & yes & yes & yes & yes \\
&Bradley-Terry consistency & yes & yes & yes & yes & yes & yes \\
&canonicity & yes & no & yes & yes & no & yes \\
&symmetry & yes & yes & yes & yes & no & yes \\
& Bayes covariant & no & no & no & yes & yes & yes \\
\hline
\multirow{5}{*}{\begin{turn}{90}\makecell{non-exact}\end{turn}}
&Hinton's minimality & N/A & \multicolumn{5}{c}{see Section \ref{sec:exp1}} \\
&robustness & N/A & \multicolumn{5}{c}{see Section \ref{sec:exp1}} \\
&lack of parameters & yes & yes & yes & yes & yes & yes \\
& amounts to a linear system & no & yes & yes & yes & yes & yes\\
& linear system of special form & N/A & no & yes & yes & yes & yes \\
& iterative method & yes & yes & yes & no & not needed & not needed\\
\hline
\end{tabular}
\caption{Summary of properties of coupling methods described in Section \ref{sec:coupling}. The method of Hastie-Tibshirani is omitted from evaluation in Section \ref{sec:exp1}, because earlier evaluation showed it lacks robustness compared to the method of Wu-Lin-Weng [\cite{wu2004probability}].}
\label{tab:summaryCoupling}
\end{table}

>>>>

A notable example which uses this principle is the coupling method introduced in \cite{hastie1998classification}. Their method uses Kullback-Leibler divergence $\divm_\textrm{KL}$
\begin{align*}
	\divm\nolimits_\textrm{KL} (\boldsymbol{p}, \boldsymbol{q})= \sum p_i \log (p_i/ q_i)
\end{align*}

Kullback-Leibler divergence strongly penalizes  mispredictions when the true class is predicted to have very low probability. In the context of classification, when we are primarily concerned with accuracy , one could advantageously use other Bregman divergence derived from a proper score \cite{gneiting2007strictly, buja2005loss}.

For instance, a popular alternative is divergence arising from the quadratic  score \cite{gneiting2007strictly} (also known as Brier score \cite{brier1950verification}). This divergence takes the form
$$
\divm\nolimits_\textrm{quad} (\boldsymbol{p}, \boldsymbol{q})= \sum (p_i - q_i)^2.
$$
Unlike Kullback-Leibler divergence, the penalty for incorrect prediction with arbitrarily low probability is bounded (by one). This quadratic score is the basis of the popular Wu-Lin-Weng's method \cite{wu2004probability}.

@@

We note that the method is symmetric. The choice of quadratic scoring rule can also be viewed as canonical in view of axiomatic characterization given in  \cite{selten1998axiomatic}.  However the choice of weights $w_{ij}$ to make the method computationally amenable is somewhat arbitrary. Thus we do not view Wu-Lin-Weng's method as canonical.

@@

As an example, consider the classification problem with three classes. We have three arboreal classifiers $\boldsymbol{s}_1, \boldsymbol{s}_2, \boldsymbol{s}_3$ which correspond to star graphs centered respectively at classes $C_1, C_2, C_3$. The classifier 
\begin{align}
\boldsymbol{e}_3 \stackrel{def}{=} \frac13 \boldsymbol{s}_1 \oplus \frac13 \boldsymbol{s}_2 \oplus \frac 13 \boldsymbol{s}_3  \label{eq:normal3}
\end{align}
is Bayes covariant \emph{and} symmetric. A key result proved in \cite{vsuch2016bayes} is that it is a unique  classifier having these two properties.

@@

\begin{table}[!ht]
\begin{tabular}{cm{2.5cm}m{1.5cm}m{1.5cm}ccm{1.5cm}m{1.5cm}}
&property & Hastie-Tibshirani & Wu-Lin-Weng & radial & normal & arboreal & Hinton's oracle \\
\hline 
\multirow{5}{*}{\begin{turn}{90}\makecell{exact}\end{turn}}
&uniqueness &  yes & yes & yes & yes & yes & yes \\
&Bradley-Terry consistency & yes & yes & yes & yes & yes & yes \\
&canonicity & yes & no & yes & yes & no & yes \\
&symmetry & yes & yes & yes & yes & no & yes \\
& Bayes covariant & no & no & no & yes & yes & yes \\
\hline
\multirow{5}{*}{\begin{turn}{90}\makecell{non-exact}\end{turn}}
&Hinton's minimality & N/A & \multicolumn{5}{c}{see Section \ref{sec:exp1}} \\
&robustness & N/A & \multicolumn{5}{c}{see Section \ref{sec:exp1}} \\
&lack of parameters & yes & yes & yes & yes & yes & yes \\
& amounts to a linear system & no & yes & yes & yes & yes & yes\\
& linear system of special form & N/A & no & yes & yes & yes & yes \\
& iterative method & yes & yes & yes & no & not needed & not needed\\
\hline
\end{tabular}
\caption{Summary of properties of coupling methods described in Section \ref{sec:coupling}. The method of Hastie-Tibshirani is omitted from evaluation in Section \ref{sec:exp1}, because earlier evaluation showed it lacks robustness compared to the method of Wu-Lin-Weng \cite{wu2004probability}.}
\label{tab:summaryCoupling}
\end{table}
